---
title: "Categorization of variables and descriptive tables"
author: "Núria Peñuelas Peñarroya"
date: "2024-11-25"
output: html_document
---

# **NORMALITY TESTS**
#We perform Anderson-Darling tests to examine the distribution of the numerical variables

```{r packages}
library(gtsummary)
library(magrittr)
library(nortest)
library(knitr)
library(dplyr)
library(writexl)
```

# Normality in continuous variables, we choose Ad test because we have a datset n>50. For larger datasets (n>50), other tests like the Kolmogorov-Smirnov test, Anderson-Darling test, or Lilliefors test may be more robust alternatives. 

```{r}
numeric_data <- select_if(TP53status_clinical, is.numeric)
print(colnames(numeric_data))


# Filter numeric columns with sufficient unique values
numeric_data <- numeric_data %>% select_if(~ length(unique(.)) > 1)

# Remove rows with NA in numeric data
cleaned_data <- numeric_data %>% drop_na()

# Create an empty dataframe to store results
results_df <- data.frame(Variable = character(),
                         Test_utilizado = character(),
                         Valor_p = numeric(),
                         Interpretación = character(),
                         stringsAsFactors = FALSE)

# Loop to check normality
for (col in colnames(cleaned_data)) {
  # Extract column data
  data_to_test <- cleaned_data[[col]]
  
  # Only proceed if there are enough data points
  if (length(data_to_test) > 1) {
    # Test Anderson-Darling
    ad_test <- ad.test(data_to_test)
    
    # Interpretation
    if (!is.na(ad_test$p.value) && ad_test$p.value < 0.05) {
      normality <- "Non-normal distribution"
    } else if (!is.na(ad_test$p.value)) {
      normality <- "Normal distribution"
    } else {
      normality <- "No p-value"
    }
    
    # Add the results into the dataframe
    results_df <- rbind(results_df, data.frame(Variable = col,
                                               Test_utilizado = "Anderson-Darling",
                                               Valor_p = ifelse(is.na(ad_test$p.value), NA, round(ad_test$p.value, digits = 3)),
                                               Interpretación = normality))
  }
}

# Print results
print(results_df)

#We print the results
knitr::kable(results_df, caption = "Summary normal distribution of numeric variables")

```


#As we see that all variables are non-normal, we will perform Mann–Whitney U test in all the comparisons for numeric variables. Mann–Whitney U test, also called Mann–Whitney–Wilcoxon (MWW/MWU), Wilcoxon rank-sum test, or Wilcoxon–Mann–Whitney test is a nonparametric test appropiate for non-normal variables. For tables where there are 3 groups, we will use kruskal wallis for numeric variables.


# **DESCRIPTIVE TABLES**

```{r}
#Select the variables of interest
descriptive_HTG<- TP53status_clinical %>% select(
    "hpv_status",
    "age",
    "figo_ini_ava",
    "tumor_size",
    "tumor_size_category",
    "depth_invasion",
    "depth_invasion_category",
    "hist_type_category",
    "vli",
    "nodal_status",
    "margins",
    "inflammatory_lesions_group",
    "radiotherapy",
    "surgery_type_category",
    "tp53_status_combined",
    "p53_ihc_status",
    "death_disease_60m",
    "recurrence_60m"
    )
```


#1. By HPV status
```{r}
# Ensure hpv_status is a factor (if not already)
descriptive_HTG$hpv_status <- as.factor(descriptive_HTG$hpv_status)


descriptive_table_HPVstatus <- descriptive_HTG %>%
  tbl_summary(
    by = hpv_status,
    statistic = list(
      all_categorical() ~ "{n} / {N} ({p}%)",
      all_continuous() ~ "{median} ({p25}, {p75})"
    ),
    type = list(
      figo_ini_ava ~ "categorical",
      tumor_size_category ~ "categorical",
      depth_invasion_category ~ "categorical",
      hist_type_category ~ "categorical",
      vli ~ "categorical",
      nodal_status ~ "categorical",
      margins ~ "categorical",
      radiotherapy ~ "categorical",
      surgery_type_category ~ "categorical",
      tp53_status_combined ~ "categorical",
      p53_ihc_status ~ "categorical",
      recurrence_60m ~ "categorical",
      death_disease_60m ~ "categorical"
    ),
    digits = list(
      all_categorical() ~ c(0, 0, 1),
      all_continuous() ~ 1
    )
  ) %>%
  add_p(
    test = list(
      all_categorical() ~ function(data, variable, by, ...) {
        # Fisher test with simulated p-value
        result <- fisher.test(data[[variable]], data[[by]], simulate.p.value = TRUE)
        tibble::tibble(p.value = result$p.value)  # Return as tibble
      },
      all_continuous() ~ function(data, variable, by, ...) {
        # Wilcoxon test with approximate p-value
        result <- wilcox.test(data[[variable]] ~ data[[by]], exact = FALSE)
        tibble::tibble(p.value = result$p.value)  # Return as tibble
      }
    ),
    pvalue_fun = ~ style_pvalue(.x, digits = 3)
  )

descriptive_table_HPVstatus

```






#2. By C2 and C3
#At the begining clusters were named C1(mostly HPV-independent, n=14) C2(mostly HPV-independent, n=42) C3 (mostly HPV-associated) and now we have changed the naming of C3 to C1 and C1 to C3 to keepn on with the logic of the paper


```{r}
#Select the variables of interest
descriptive_HTG<- TP53status_clinical %>% select(
    "hpv_status",
    "Clusters",
    "age",
    "figo_ini_ava",
    "tumor_size",
    "tumor_size_category",
    "depth_invasion",
    "depth_invasion_category",
    "hist_type_category",
    "vli",
    "nodal_status",
    "margins",
    "inflammatory_lesions_group",
    "radiotherapy",
    "surgery_type_category",
    "tp53_status_combined",
    "p53_ihc_status",
    "death_disease_60m",
    "recurrence_60m"
    )
```

```{r}
descriptive_table_2clusters <- descriptive_HTG %>%
  filter(!(Clusters== "Cluster 3")) %>%  # Exclude rows where cluster is "Cluster 3"
  droplevels() %>%  # Drop the unused "Cluster 3" factor level
  tbl_summary(
    by = Clusters,
    statistic = list(
      all_categorical() ~ "{n} / {N} ({p}%)",
      all_continuous() ~ "{median} ({p25}, {p75})"
    ),
    type = list(
      figo_ini_ava ~ "categorical",
      hpv_status ~ "categorical",
      tumor_size_category ~ "categorical",
      depth_invasion_category ~ "categorical",
      hist_type_category ~ "categorical",
      inflammatory_lesions_group ~ "categorical",
      vli ~ "categorical",
      nodal_status ~ "categorical",
      margins ~ "categorical",
      radiotherapy ~ "categorical",
      surgery_type_category ~ "categorical",
      p53_ihc_status ~ "categorical",
      tp53_status_combined ~ "categorical",
      recurrence_60m ~ "categorical",
      death_disease_60m ~ "categorical"
    ),
    digits = list(
      all_categorical() ~ c(0, 0, 1),
      all_continuous() ~ 1
    )
  ) %>%
  add_p(
    test = list(
      all_categorical() ~ function(data, variable, by, ...) {
        # Fisher test with simulated p-value
        result <- fisher.test(data[[variable]], data[[by]], simulate.p.value = TRUE)
        tibble::tibble(p.value = result$p.value)  # Return as tibble
      },
      all_continuous() ~ function(data, variable, by, ...) {
        # Wilcoxon test with approximate p-value
        result <- wilcox.test(data[[variable]] ~ data[[by]], exact = FALSE)
        tibble::tibble(p.value = result$p.value)  # Return as tibble
      }
    ),
    pvalue_fun = ~ style_pvalue(.x, digits = 3)
  )

descriptive_table_2clusters
```






#3. By TP53 mutation (only HPV-indepenent cases) (Supplementary)
```{r}
descriptive_table_TP53_HPV_ind <- descriptive_HTG %>%
  filter(hpv_status == "Negative") %>%  # Filter for HPV positive
  tbl_summary(
    by = tp53_status_combined,
    statistic = list(
      all_categorical() ~ "{n} / {N} ({p}%)",
      all_continuous() ~ "{median} ({p25}, {p75})"
    ),
    type = list(
      figo_ini_ava ~ "categorical",
      tumor_size_category ~ "categorical",
      depth_invasion_category ~ "categorical",
      hist_type_category ~ "categorical",
      inflammatory_lesions_group ~ "categorical",
      vli ~ "categorical",
      nodal_status ~ "categorical",
      margins ~ "categorical",
      radiotherapy ~ "categorical",
      surgery_type_category ~ "categorical",
      p53_ihc_status ~ "categorical",
      recurrence_60m ~ "categorical",
      death_disease_60m ~ "categorical"
    ),
    digits = list(
      all_categorical() ~ c(0, 0, 1),
      all_continuous() ~ 1
    )
  ) %>%
  add_p(
    test = list(
      all_categorical() ~ function(data, variable, by, ...) {
        # Fisher test with simulated p-value
        result <- fisher.test(data[[variable]], data[[by]], simulate.p.value = TRUE)
        tibble::tibble(p.value = result$p.value)  # Return as tibble
      },
      all_continuous() ~ function(data, variable, by, ...) {
        # Wilcoxon test with approximate p-value
        result <- wilcox.test(data[[variable]] ~ data[[by]], exact = FALSE)
        tibble::tibble(p.value = result$p.value)  # Return as tibble
      }
    ),
    pvalue_fun = ~ style_pvalue(.x, digits = 3)
  )

descriptive_table_TP53_HPV_ind
```

