---
title: "Loading clinical data from Redcap"
author: "Núria Peñuelas Peñarroya"
date: "2025-01-13"
output: html_document
---

#Packages
```{r packages}
library(readxl)
library(dplyr)
library(ggplot2)
library(ggrepel)
library(tidyr)
library(redcapAPI)
library(gtsummary)
library(magrittr)
library(survival)
```

#Connection to Redcap API
```{r}
source("tokens.R")

rcon_VSCC = redcapConnection(url = api_url,
                                   token = token_VSCC)
```

#Export Redcap data by fields
```{r}
oncomine_followup <- exportRecordsTyped(
  rcon = rcon_VSCC,
  fields = c(
    "study_id",
    "age",
    "group",
    "infl_lesions",
    "figo_2021_stage",
    "figo_ini_ava",
    "tumor_size",
    "depth_invasion",
    "hist_type",
    "vli",
    "snl_result",
    "nodal_status",
    "margins",
    "radiotherapy",
    "surgery_type",
    "hpv_status",
    "ciclina_d1",
    "p53_ihc_status",
    "oncomine_analysis",
    "valid_oncomine",
    "tp53_status_onc",
    "oncomine_varclasf_1",
    "ccnd1_cnv_onc",
    "surgery_date",
    "recurrence",
    "recurrence_date",
    "last_follow_up",
    "status_last_followup",
    "valid_prognostic"
  )
)

WES <- exportRecordsTyped(
  rcon = rcon_VSCC,
  fields = c(
    "tp53_status",
    "vscc_vaf_1",
    "bioinf_analysis",
    "cnv",
    "vscc_varclasf_1",
    "ccnd1_status_new"))

```

#Merge the two different datasets from different worksheets and create a common TP53 status varibale from WES and Oncomine
```{r}
# Perform a full join to combine all rows, matching by `study_id`
merged_data <- full_join(oncomine_followup, WES, by = "study_id")

# Combine duplicate rows for the same `study_id` into a single row
# Use summarization to keep all unique non-NA values for each column
collapsed_data <- merged_data %>%
  group_by(study_id) %>%
  summarize(across(everything(), ~ paste(unique(na.omit(.)), collapse = "; "), .names = "{col}"), .groups = "drop")

TP53status_clinical<-collapsed_data

#Identify empty strings correctly as NA
TP53status_clinical$tp53_status[TP53status_clinical$tp53_status == ""] <- NA
TP53status_clinical$tp53_status_onc[TP53status_clinical$tp53_status_onc == ""] <- NA

#Check NAs
summary(TP53status_clinical$tp53_status_onc)
table(is.na(TP53status_clinical$tp53_status_onc))

summary(TP53status_clinical$tp53_status)
table(is.na(TP53status_clinical$tp53_status))

#Create in a new variable (tp53_status_combined) that if TP53 WES has no info, we consider the info from oncomine
TP53status_clinical$tp53_status_combined <- ifelse(
  is.na(TP53status_clinical$tp53_status),
  TP53status_clinical$tp53_status_onc,
  TP53status_clinical$tp53_status
)

#Check the values
summary(TP53status_clinical$tp53_status_combined)
table(is.na(TP53status_clinical$tp53_status_combined))

```

#Make IDs consistent between expression data and clinical data
```{r}
#To make ID formats consistent between clinical data and counts, we remove the first two characters after the "B" from the Study ID column in TP53status
TP53status_clinical <- TP53status_clinical %>%
  mutate(`study_id` = sub("^B\\d{2}", "B", `study_id`)) # Update Study ID in place

#Also we change these 2 cases that have an alternative study_id: B2015-15290  for B17-38865 and B2008-08482 for B13-31747, which are the ones that are in the counts_filtered_data
TP53status_clinical <- TP53status_clinical %>%
  mutate(study_id = case_when(
    study_id == "B15-15290" ~ "B17-38865",
    study_id == "B08-08482" ~ "B13-31747",
    TRUE ~ study_id  # Keep the rest unchanged
  ))
```



#Load info on HE EMT patterns: tummor budding and patterns 
```{r}
#We load pattern/tumor budding information
EMT_markers<- read_excel("EMT-Tumorbudding/20250227_Robert_combination_budding_morpho.xlsx")

# Select only the desired columns from EMT_markers
EMT_markers_selected <- EMT_markers %>% 
  dplyr::select(id,Tumor_budding, Pattern, Combination, Combination2, Inflammatory)

# Merge the selected columns with TP53status_clinical by id and Study_ID
TP53status_clinical <- dplyr::left_join(TP53status_clinical, EMT_markers_selected, 
                                   by = c("study_id" = "id"))

# View the structure of the merged dataset
str(TP53status_clinical)

#20250512_We change the name of the patterns to spray and non-spray
TP53status_clinical$Pattern <- ifelse(
  TP53status_clinical$Pattern == "spray-like",
  "spray",
  "non-spray"
)
```


#We exclude the cases that are in the clinical file but not in the expression dataset (HTG dataset)
```{r}
#Load the counts
load("~/Projects/14. Clusters Lorena Survival/Survival Clusters Lorena/counts_filtered_filtered.RData")
HTGcases<-c(colnames(counts_filtered_filtered))

# Find matching IDs
matching_ids <- intersect(HTGcases, TP53status_clinical$`study_id`)

# Print the number of matching IDs
print(length(matching_ids))  # This will give you the count of matching IDs

# Filter rows where the processed ID matches IDs in TP53status_clinical
TP53status_clinical <- TP53status_clinical %>%
  filter(`study_id` %in% matching_ids)

# Print the resulting filtered data
print(TP53status_clinical)

#Now we have the same 75 cases in both the TP53status and clinical data as in counts
```



#We create new variables in our clinical dataset
```{r}
TP53status_clinical <- TP53status_clinical %>%
  rename(group_premalign = group)

TP53status_clinical <- TP53status_clinical %>%
  mutate(group_premalign_category = gsub("VSCC \\+ DEVIL|VSCC \\+ VAAD", 
                                         "VSCC + DEVIL/VAAD", 
                                         group_premalign))


TP53status_clinical$inflammatory_lesions_group <- ifelse(
  TP53status_clinical$infl_lesions___no == "Checked", "No",
  ifelse(
    TP53status_clinical$infl_lesions___lsc == "Checked" | 
    TP53status_clinical$infl_lesions___les == "Checked" | 
    TP53status_clinical$infl_lesions___lp == "Checked", 
    "Yes", 
    NA
  )
)  
```



#Format and create new variables in TP53status_clinical
```{r}
# Create a new column Recurrence_01 in AnnotData
TP53status_clinical <- TP53status_clinical %>%
  mutate(
    # Convert numeric variables to numeric
    age = as.numeric(age),
    tumor_size = as.numeric(tumor_size),
    depth_invasion = as.numeric(depth_invasion),
    ciclina_d1= as.numeric(ciclina_d1), #There are 2 samples with no info -->NA introduced
    
     # Convert recurrence_date and surgery_date to Date type
    recurrence_date = as.Date(recurrence_date, format = "%Y-%m-%d"),  # Adjust the format if necessary
    surgery_date = as.Date(surgery_date, format = "%Y-%m-%d"),        # Adjust the format if necessary
    last_follow_up = as.Date(last_follow_up, format = "%Y-%m-%d"),  # Adjust the format if necessary

    
    recurrence = ifelse(recurrence == "Yes", 1, 
                        ifelse(recurrence == "No", 0, NA)),
    #death_disease = ifelse(status_last_followup %in% c("Dead of disease", "Dead of other causes"), 1, 0),
    death_disease = ifelse(status_last_followup == "Dead of disease", 1, 0),
    
    time_recurrence = as.numeric(difftime(recurrence_date, surgery_date, 
                                          units = 'days') / (365.25 / 12)),
    
    time_last_followup = as.numeric(difftime(last_follow_up, surgery_date, 
                                             units = 'days') / (365.25 / 12)),
    
    time_death_surv = time_last_followup,
    
    time_recurrence_surv = ifelse(!is.na(time_recurrence), time_recurrence, time_last_followup)
  )
```


#We exclude those with no TP53information in the clinical file and in the counts as well
```{r}
TP53status_clinical <- TP53status_clinical[!TP53status_clinical$study_id %in% c("B12-27987", "B16-24861"), ]

counts_filtered_filtered <- counts_filtered_filtered[, !(names(counts_filtered_filtered) %in% c("B12-27987", "B16-24861"))]

#15/05/2025: There is another that has no info on TP53! B18-21625!!! De moment només el trec per fer els analisis de survival ajustant per TP53...no als counts...
TP53status_clinical <- TP53status_clinical[!is.na(TP53status_clinical$tp53_status_combined), ]

#19/05/2025: La Lorena m'ha dit que ella té el cas  B18-21625 com a wt..
TP53status_clinical$tp53_status_combined[TP53status_clinical$study_id == "B18-21625"] <- "Wt"


# Check results
dim(counts_filtered_filtered)
#There are now n=73 (75 - 2 cases without no TP53status information) HPV-Independent, 10513 genes

#We check the numbers of TP53 mutated cases in HPV-independent group
#table(TP53status_clinical$group)
#There are n=42 TP53-mutated and n=12 TP53-wt in HPV-independent cases


```
#Format and create new variables in our clinical dataset
```{r}
#SNL result
#There are 30 cases whith no info on SNL result, but they are not NA...
table(TP53status_clinical$snl_result)
is.na(TP53status_clinical$snl_result)

#We consider negative when no info
TP53status_clinical <- TP53status_clinical %>%
  mutate(snl_result = ifelse(!snl_result %in% c("Negative", "Positive"), "Negative", snl_result))

#Nodal status
#There are 10 cases whith nd in nodal status and 1 with no info, but no NA
table(TP53status_clinical$nodal_status)
is.na(TP53status_clinical$snl_result)

# Replace all "nd" values with 0
TP53status_clinical$nodal_status[TP53status_clinical$nodal_status == "nd"] <- "0"
table(TP53status_clinical$nodal_status)

#We check the case that has an empty string
print(TP53status_clinical$nodal_status[TP53status_clinical$study_id == "B23-26785"])

# Update nodal_status to "0" for the specific id
TP53status_clinical$nodal_status[TP53status_clinical$study_id == "B23-26785"] <- "0"

# Verify the update
print(TP53status_clinical$nodal_status[TP53status_clinical$study_id == "B23-26785"])


#Categorize some categorical/continuous variables to two categories
TP53status_clinical <- TP53status_clinical %>%
  mutate(
    hpv_status= as.factor(hpv_status),
    hist_type_category = case_when(
      hist_type %in% c("Basaloid", "Warty", "Non-keratinizing",
                       "Basaloid/Warty","Basaloid/Keratinizing","Warty/Keratinizing") ~
                       "Basaloid/warty/nonkeratinizing",
      hist_type %in% c("Keratinizing", "Verrucous") ~ "Keratinizing/verrucous",
      TRUE ~ "Unknown" # For unexpected or empty values
    ),
    tumor_size_category = case_when(
      tumor_size <= 25 ~ "≤25 mm",
      tumor_size > 25 ~ ">25 mm",
      TRUE ~ "Unknown"  # Handles missing or unexpected values
    ),
    depth_invasion_category = case_when(
      depth_invasion <= 5 ~ "≤5 mm",
      depth_invasion > 5 ~ ">5 mm",
      TRUE ~ "Unknown"  # Handles missing or unexpected values
    ),
    surgery_type_category = case_when(
      surgery_type == "Local exeresis" ~ "Local wide excision",
      surgery_type == "Partial vulvectomy" ~ "Local wide excision",
      surgery_type == "Radical vulvectomy" ~ "Radical vulvectomy",
      TRUE ~ surgery_type  # Keeps other values unchanged if any
    ),
    vli=as.factor(vli))
```



# Convert some variables to factors and explicitly set reference levels for cox analysis
```{r}
# Convert to factor and explicitly set reference levels for cox analysis
TP53status_clinical$hpv_status <- factor(TP53status_clinical$hpv_status, levels = c("Positive", "Negative"))
TP53status_clinical$figo_ini_ava <- factor(TP53status_clinical$figo_ini_ava, levels = c("0 (Initial)", "1 (Advanced)"))
TP53status_clinical$margins <- factor(TP53status_clinical$margins, levels = c("Free", "Affected"))
TP53status_clinical$nodal_status <- factor(TP53status_clinical$nodal_status, levels = c("0", "1"))

TP53status_clinical$tumor_size_category <- factor(
  TP53status_clinical$tumor_size_category,
  levels = c("≤25 mm", ">25 mm")
)
TP53status_clinical$depth_invasion_category <- factor(
  TP53status_clinical$depth_invasion_category,
  levels = c("≤5 mm", ">5 mm")
)

TP53status_clinical$hist_type_category <- factor(
  TP53status_clinical$hist_type_category,
  levels = c("Basaloid/warty/nonkeratinizing", "Keratinizing/verrucous")
)
```



#We check that we have the same cases in our clinical file and in the expression dataset (HTG counts)
```{r}
# Find matching IDs
matching_ids <- intersect(HTGcases, TP53status_clinical$`study_id`)
print(matching_ids)


# Find IDs in HTGcases that are not in TP53status_clinical$study_id
nonmatching_in_HTGcases <- setdiff(HTGcases, TP53status_clinical$study_id)
print("Nonmatching IDs in HTGcases:")
print(nonmatching_in_HTGcases)
```




#We prepare the data for HTG function
```{r}
identical(colnames(counts_filtered_filtered), TP53status_clinical$study_id)
#We see that the colnames in counts and $study_id in TP53status_clinical are not identical and they have to be for TME function
```

```{r}
#HTG function needs a data.frame
TP53status_clinical<- as.data.frame(TP53status_clinical)

#HTG function looks for "id" column in Annotdata:
TP53status_clinical <- TP53status_clinical %>%
  rename(id = `study_id`)

## HTG function requires rownames and columns to be ordered in the same 
TP53status_clinical <- TP53status_clinical[order(TP53status_clinical$id), ]
counts_filtered_filtered<- counts_filtered_filtered[,order(colnames(counts_filtered_filtered))]

#We confirm that the rows in clinical file correspond to the columns in the counts file
identical(colnames(counts_filtered_filtered), TP53status_clinical$id)
```


