---
title: "Clusters Manhatan-Ward2"
author: "Núria Peñuelas Peñarroya"
date: "2025-01-16"
output: html_document
---

```{r packages}
library(survminer)
library(survival)
library(coxphf)
library(knitr)
library(readr)
```



#There is one case that has been just very recently updated to invalid prognosis(B2018-05127) but for neoadjuvant therapy, not palliative treatment so we still want to include it.
```{r}
# Manually include B2018-05127 even if marked as "No"
valid_ids <- TP53status_clinical$id[
  TP53status_clinical$valid_prognostic == "Yes" |
  TP53status_clinical$id == "B18-05127"
]

# Filter the clinical dataframe accordingly
TP53status_clinical <- TP53status_clinical[TP53status_clinical$id %in% valid_ids, ]

# Filter the RNA expression dataframe to keep only valid columns
counts_filtered_filtered <- counts_filtered_filtered[, colnames(counts_filtered_filtered) %in% valid_ids]

# Verify dimensions
print(dim(TP53status_clinical))       # Should now include B2018-05127
print(dim(counts_filtered_filtered))  # Should match updated sample IDs

```

#We load information on EMT pattern by IHC Cadherin and Vimentin and IHC mTORc results
```{r}

EMT_VIM_Cadh_mTORC_info <- read_csv("EMT_VIM+Cadh_mTORC_info.csv")



# We count how many TP53 IDs are present in the EMT dataset, to confirm we have info on all 70
matching_ids <- TP53status_clinical$id %in% EMT_VIM_Cadh_mTORC_info$Tumor_Sample_ID
sum(matching_ids)
#We have info on all 70 cases

# View non-matching IDs
unmatched_ids <- TP53status_clinical$id[!matching_ids]
unmatched_ids


# Subset relevant columns from EMT_VIM_Cadh_mTORC_info
subset_info <- EMT_VIM_Cadh_mTORC_info[,c("Tumor_Sample_ID","Clusters","EMT_phenotype_Ecadh_Vimentin", "mTOR")]


# Rename columns to make merge clearer
colnames(subset_info)[colnames(subset_info) == "Tumor_Sample_ID"] <- "id"

# Merge with TP53status_clinical on 'id'
TP53status_clinical<- merge(TP53status_clinical, subset_info, by = "id", all.x = TRUE)
```


#We create survival variables for different cutoffs
#1. All 3 clusters
```{r}

# Define the time points for follow-up
time_points <- c(10, 12, 20, 24, 30, 36, 40, 48, 50, 60)

# Loop through each cutoff we want to assess to create time-restricted variables
for (time_point in time_points) {
  TP53status_clinical <- TP53status_clinical %>%
    mutate(
      # Create time-restricted survival variables: cap the times at the time_point
      !!paste0("time_recurrence_surv_", time_point, "m") := pmin(time_recurrence_surv, time_point),  # cap to time_point
      !!paste0("time_death_surv_", time_point, "m") := pmin(time_death_surv, time_point),  # cap to time_point

      # Create event indicator variables for recurrence and death
      # If the event time is greater than the time point, event does not occur (0)
      !!paste0("recurrence_", time_point, "m") := ifelse(time_recurrence_surv <= time_point & recurrence == 1, 1, 0),
      !!paste0("death_disease_", time_point, "m") := ifelse(time_death_surv <= time_point & death_disease == 1, 1, 0)
    )
}

# We convert the main variables (recurrence and death disease) as a factor
TP53status_clinical <- TP53status_clinical %>%
  mutate(
    # Convert the base recurrence and death variables to factors
    recurrence = factor(recurrence, levels = c(0, 1), labels = c("No", "Yes")),
    death_disease = factor(death_disease, levels = c(0, 1), labels = c("No", "Yes"))
  )

# Loop through each time point to convert the event indicators into factors
for (time_point in time_points) {
  TP53status_clinical <- TP53status_clinical %>%
    mutate(
      !!paste0("recurrence_", time_point, "m") := factor(get(paste0("recurrence_", time_point, "m")), levels = c(0, 1), labels = c("No", "Yes")),
      !!paste0("death_disease_", time_point, "m") := factor(get(paste0("death_disease_", time_point, "m")), levels = c(0, 1), labels = c("No", "Yes"))
    )
}

# Verify structure
str(TP53status_clinical)

```


# We check for proper conversion of all variables
```{r}
# Check missing values for critical variables
colSums(is.na(TP53status_clinical[c("time_last_followup", "time_recurrence", "recurrence", "death_disease")]))

summary(TP53status_clinical$time_recurrence_surv)
summary(TP53status_clinical$time_death_surv)


for (time_point in time_points) {
  cat("\nTime Point:", time_point, "months\n")
  print(table(TP53status_clinical[[paste0("recurrence_", time_point, "m")]]))
}

for (time_point in time_points) {
  TP53status_clinical <- TP53status_clinical %>%
    mutate(
      time_recurrence_surv_temp = pmin(time_recurrence_surv, time_point),
      recurrence_temp = ifelse(time_recurrence_surv <= time_point & recurrence == 1, 1, 0)
    )
  cat("\nTime Point:", time_point, "\n")
  print(table(TP53status_clinical$recurrence_temp, useNA = "ifany"))
}
```




#1. By Clusters
#Pattern_Kaplan Meyer recurrence individual plot
```{r}
# Define the custom colors with more reddish orange
custom_colors <- c("#009E73", "#D45F00", "#0072B2")

# Filter for Cluster 1 and 2 only
filtered_data <- TP53status_clinical %>%
  filter(Clusters %in% c("Cluster 1", "Cluster 2"))
dim(filtered_data)
#We have 57 cases from C1+C2

# Create the survival object
surv_obj <- Surv(time = filtered_data$time_recurrence_surv_60m, 
                 event = as.numeric(filtered_data$recurrence_60m == "Yes"))

# Fit the Kaplan-Meier survival model
km_fit <- survfit(surv_obj ~ Clusters, data = filtered_data)

# Plot the Kaplan-Meier survival curves
p <- ggsurvplot(
  km_fit, 
  data = filtered_data,
  pval = TRUE,
  title = "Kaplan-Meier Survival Curve by Clusters_Recurrence (Cluster 1 & 2 Only)",
  risk.table = TRUE,
  risk.table.height = 0.3,
  xlab = "Time to Event (Months)",
  ylab = "Survival Probability",
  break.time.by = 10,
  palette = custom_colors,
  legend = c(0.8, 0.2),
  legend.title = "Clusters"
)

# Display the plot
print(p)

```



#Pattern_Kaplan Meyer death individual plot
```{r}
# Create the survival object
surv_obj <- Surv(time = filtered_data$time_death_surv_60m, 
                 event = as.numeric(filtered_data$death_disease_60m == "Yes"))

# Fit the Kaplan-Meier survival model
km_fit <- survfit(surv_obj  ~ Clusters, data = filtered_data)


# Plot the Kaplan-Meier survival curves
p <- ggsurvplot(
  km_fit, 
  data = filtered_data,
  pval = TRUE,                    # Display the p-value for the log-rank test
  title = "Kaplan-Meier Survival Curve by Clusters_Death",
  risk.table = TRUE,              # Include the risk table
  risk.table.height = 0.3,        # Adjust the height of the risk table
  xlab = "Time to Event (Months)",# Label for x-axis
  ylab = "Survival Probability",   # Label for y-axis
  break.time.by = 10,               # Adjust time interval to match plot
  palette = custom_colors, # Use the custom color palette
  legend = c(0.8, 0.2),           # Adjust legend position (optional)
  legend.title = "Clusters"  # Customize legend labels
) 

# Add the title to the plot
p$plot <- p$plot + ggtitle("Kaplan-Meier Survival Curve by Clusters_Death")

# Display the plot
print(p)


```




#2. By EMT pattern Cadh-Vim
#Kaplan Meyer recurrence individual plot
```{r}



# Filter out rows where EMT_phenotype_Ecadh_Vimentin is "N/A"
filtered_data <- TP53status_clinical %>%
  filter(Clusters %in% c("Cluster 1", "Cluster 2"))%>%
  filter(EMT_phenotype_Ecadh_Vimentin != "N/A")

dim(filtered_data)
#We have 54 cases with info on EMT (and C1 and C2)

# Create the survival object
surv_obj <- Surv(time = filtered_data$time_recurrence_surv_60m, 
                 event = as.numeric(filtered_data$recurrence_60m == "Yes"))

# Fit the Kaplan-Meier survival model
km_fit <- survfit(surv_obj ~ EMT_phenotype_Ecadh_Vimentin, data = filtered_data)

# Plot the Kaplan-Meier survival curves
p <- ggsurvplot(
  km_fit, 
  data = filtered_data,
  pval = TRUE,
  title = "Kaplan-Meier Survival Curve by EMT_phenotype_Ecadh_Vimentin_Recurrence (Cluster 1 & 2 Only)",
  risk.table = TRUE,
  risk.table.height = 0.3,
  xlab = "Time to Event (Months)",
  ylab = "Survival Probability",
  break.time.by = 10,
  palette = custom_colors,
  legend = c(0.8, 0.2),
  legend.title = "Clusters"
)

# Display the plot
print(p)

```



#Pattern_Kaplan Meyer death individual plot
```{r}
# Create the survival object
surv_obj <- Surv(time = filtered_data$time_death_surv_60m, 
                 event = as.numeric(filtered_data$death_disease_60m == "Yes"))

# Fit the Kaplan-Meier survival model
km_fit <- survfit(surv_obj  ~ EMT_phenotype_Ecadh_Vimentin, data = filtered_data)


# Plot the Kaplan-Meier survival curves
p <- ggsurvplot(
  km_fit, 
  data = filtered_data,
  pval = TRUE,                    # Display the p-value for the log-rank test
  title = "Kaplan-Meier Survival Curve by Clusters_Death",
  risk.table = TRUE,              # Include the risk table
  risk.table.height = 0.3,        # Adjust the height of the risk table
  xlab = "Time to Event (Months)",# Label for x-axis
  ylab = "Survival Probability",   # Label for y-axis
  break.time.by = 10,               # Adjust time interval to match plot
  palette = custom_colors, # Use the custom color palette
  legend = c(0.8, 0.2),           # Adjust legend position (optional)
  legend.title = "Clusters"  # Customize legend labels
) 

# Add the title to the plot
p$plot <- p$plot + ggtitle("Kaplan-Meier Survival Curve by EMT_phenotype_Ecadh_Vimentin_Death (Cluster 1 & 2 Only)")

# Display the plot
print(p)
```



#3. By mTORC IHC
#Kaplan Meyer recurrence individual plot
```{r}

# Define the custom colors with more reddish orange
custom_colors <- c("#009E73", "#D45F00", "#0072B2")


# Recode mTOR values in the original data
filtered_data <- TP53status_clinical %>%
  mutate(mTOR = ifelse(mTOR %in% c("n/a (no tumor)", "no tumor", "N/A"), NA, mTOR))
dim(filtered_data)
#70 cases

# Filter for Clusters C1 and C2 and remove NAs in mTOR
filtered_data <- filtered_data %>%
  filter(Clusters %in% c("Cluster 1", "Cluster 2") & !is.na(mTOR))

dim(filtered_data)
#We have 51 cases with info mTOR (and C1 and C2)

# Create the survival object
surv_obj <- Surv(time = filtered_data$time_recurrence_surv_60m, 
                 event = as.numeric(filtered_data$recurrence_60m == "Yes"))

# Fit the Kaplan-Meier survival model
km_fit <- survfit(surv_obj ~ mTOR, data = filtered_data)

# Plot the Kaplan-Meier survival curves
p <- ggsurvplot(
  km_fit, 
  data = filtered_data,
  pval = TRUE,
  title = "Kaplan-Meier Survival Curve by mTOR_Recurrence (Cluster 1 & 2 Only)",
  risk.table = TRUE,
  risk.table.height = 0.3,
  xlab = "Time to Event (Months)",
  ylab = "Survival Probability",
  break.time.by = 10,
  palette = custom_colors,
  legend = c(0.8, 0.2),
  legend.title = "Clusters"
)

# Display the plot
print(p)

```



#Pattern_Kaplan Meyer death individual plot
```{r}
# Create the survival object
surv_obj <- Surv(time = filtered_data$time_death_surv_60m, 
                 event = as.numeric(filtered_data$death_disease_60m == "Yes"))

# Fit the Kaplan-Meier survival model
km_fit <- survfit(surv_obj  ~ mTOR, data = filtered_data)


# Plot the Kaplan-Meier survival curves
p <- ggsurvplot(
  km_fit, 
  data = filtered_data,
  pval = TRUE,                    # Display the p-value for the log-rank test
  title = "Kaplan-Meier Survival Curve by mTOR_Death",
  risk.table = TRUE,              # Include the risk table
  risk.table.height = 0.3,        # Adjust the height of the risk table
  xlab = "Time to Event (Months)",# Label for x-axis
  ylab = "Survival Probability",   # Label for y-axis
  break.time.by = 10,               # Adjust time interval to match plot
  palette = custom_colors, # Use the custom color palette
  legend = c(0.8, 0.2),           # Adjust legend position (optional)
  legend.title = "Clusters"  # Customize legend labels
) 

# Add the title to the plot
p$plot <- p$plot + ggtitle("Kaplan-Meier Survival Curve by mTOR_Death (Cluster 1 & 2 Only)")

# Display the plot
print(p)
```





#4. By HE EMT Pattern

#Kaplan Meyer recurrence individual plot
```{r}
# Create the survival object
surv_obj <- Surv(time = TP53status_clinical$time_recurrence_surv_60m, 
                 event = as.numeric(TP53status_clinical$recurrence_60m == "Yes"))

# Fit the Kaplan-Meier survival model
km_fit <- survfit(surv_obj  ~ Pattern, data = TP53status_clinical)


# Plot the Kaplan-Meier survival curves
p <- ggsurvplot(
  km_fit, 
  data = TP53status_clinical,
  pval = TRUE,                    # Display the p-value for the log-rank test
  title = "Kaplan-Meier Survival Curve by Pattern_Recurrence",
  risk.table = TRUE,              # Include the risk table
  risk.table.height = 0.3,        # Adjust the height of the risk table
  xlab = "Time to Event (Months)",# Label for x-axis
  ylab = "Survival Probability",   # Label for y-axis
  break.time.by = 10,               # Adjust time interval to match plot
  palette = custom_colors, # Use the custom color palette
  legend = c(0.8, 0.2),           # Adjust legend position (optional)
  legend.title = "Clusters"  # Customize legend labels
) 

# Add the title to the plot
p$plot <- p$plot + ggtitle("Kaplan-Meier Survival Curve by Pattern_Recurrence")

# Display the plot
print(p)
```




#Pattern_Kaplan Meyer death individual plot
```{r}
# Create the survival object
surv_obj <- Surv(time = TP53status_clinical$time_death_surv_60m, 
                 event = as.numeric(TP53status_clinical$death_disease_60m == "Yes"))

# Fit the Kaplan-Meier survival model
km_fit <- survfit(surv_obj  ~ Pattern, data = TP53status_clinical)


# Plot the Kaplan-Meier survival curves
p <- ggsurvplot(
  km_fit, 
  data = TP53status_clinical,
  pval = TRUE,                    # Display the p-value for the log-rank test
  title = "Kaplan-Meier Survival Curve by Pattern_Death",
  risk.table = TRUE,              # Include the risk table
  risk.table.height = 0.3,        # Adjust the height of the risk table
  xlab = "Time to Event (Months)",# Label for x-axis
  ylab = "Survival Probability",   # Label for y-axis
  break.time.by = 10,               # Adjust time interval to match plot
  palette = custom_colors, # Use the custom color palette
  legend = c(0.8, 0.2),           # Adjust legend position (optional)
  legend.title = "Clusters"  # Customize legend labels
) 

# Add the title to the plot
p$plot <- p$plot + ggtitle("Kaplan-Meier Survival Curve by Pattern_Death")

# Display the plot
print(p)


```


#EMT HE Pattern_Cox univariate recurrence
```{r}
univ_recurrence_Pattern <- coxphf(
  Surv(time_recurrence_surv_60m, recurrence_60m) ~ Pattern,
  data = TP53status_clinical,
  maxstep = 0.0001, 
  maxit = 100000
)

# Extract coefficients, hazard ratios, and confidence intervals
univ_recurrence_Pattern_table <- data.frame(
  coef = univ_recurrence_Pattern$coefficients,
  HR = exp(univ_recurrence_Pattern$coefficients),
  lower_095 = univ_recurrence_Pattern$ci.lower,
  upper_095 = univ_recurrence_Pattern$ci.upper,
  pval = univ_recurrence_Pattern$prob
)

# Display the table
kable(univ_recurrence_Pattern_table, digits = 3)
```

#EMT HE Pattern_Cox multivariate recurrence
```{r}
# Ensure TP53 is a factor
TP53status_clinical$tp53_status_combined <- factor(TP53status_clinical$tp53_status_combined)

# Set "Wt" as the reference level
TP53status_clinical$tp53_status_combined <- relevel(TP53status_clinical$tp53_status_combined, ref = "Wt")


multiv_recurrence_Pattern <- coxphf(
  Surv(time_recurrence_surv_60m, recurrence_60m) ~ Pattern + figo_ini_ava + tp53_status_combined,
  data = TP53status_clinical,
  maxstep = 0.0001, 
  maxit = 100000
)

# Extract coefficients, hazard ratios, and confidence intervals
multiv_recurrence_Pattern_table <- data.frame(
  coef = multiv_recurrence_Pattern$coefficients,
  HR = exp(multiv_recurrence_Pattern$coefficients),
  lower_095 = multiv_recurrence_Pattern$ci.lower,
  upper_095 = multiv_recurrence_Pattern$ci.upper,
  pval = multiv_recurrence_Pattern$prob
)

# Display the table
kable(multiv_recurrence_Pattern_table, digits = 3)
```




#EMT HE Pattern_Cox univariate death
```{r, echo=FALSE}
multiv_death_Pattern <- coxphf(
  Surv(time_death_surv_60m, death_disease_60m) ~ Pattern,
  data = TP53status_clinical,
  maxstep = 0.0001, 
  maxit = 100000
)

# Extract coefficients, hazard ratios, and confidence intervals
multiv_death_Pattern_table <- data.frame(
  coef = multiv_death_Pattern$coefficients,
  HR = exp(multiv_death_Pattern$coefficients),
  lower_095 = multiv_death_Pattern$ci.lower,
  upper_095 = multiv_death_Pattern$ci.upper,
  pval = multiv_death_Pattern$prob
)

# Display the table
kable(multiv_death_Pattern_table, digits = 3)

```




#EMT HE Pattern_Cox multivariate death
```{r, echo=FALSE}
multiv_death_Pattern <- coxphf(
  Surv(time_death_surv_60m, death_disease_60m) ~ Pattern + figo_ini_ava + tp53_status_combined,
  data = TP53status_clinical,
  maxstep = 0.0001, 
  maxit = 100000
)

# Extract coefficients, hazard ratios, and confidence intervals
multiv_death_Pattern_table <- data.frame(
  coef = multiv_death_Pattern$coefficients,
  HR = exp(multiv_death_Pattern$coefficients),
  lower_095 = multiv_death_Pattern$ci.lower,
  upper_095 = multiv_death_Pattern$ci.upper,
  pval = multiv_death_Pattern$prob
)

# Display the table
kable(multiv_death_Pattern_table, digits = 3)

```




#We plot death and recurrence HR with a forst plot
```{r}
multiv_death_Pattern_table$Variable <- rownames(multiv_death_Pattern_table)
multiv_recurrence_Pattern_table$Variable <- rownames(multiv_recurrence_Pattern_table)
multiv_death_Pattern_table$Outcome <- "Death"
multiv_recurrence_Pattern_table$Outcome <- "Recurrence"
combined_table_Pattern <- rbind(multiv_death_Pattern_table, multiv_recurrence_Pattern_table)
library(ggplot2)

combined_table_Pattern$Variable <- factor(combined_table_Pattern$Variable, levels = c(
  "figo_ini_ava1 (Advanced)",
  "Patternspray",
  "tp53_status_combinedMut"
))

forestplot_recurrence_death_Pattern<-ggplot(combined_table_Pattern, aes(x = reorder(Variable, -HR), y = HR, color = Outcome)) +
  geom_point(position = position_dodge(width = 0.6), size = 3) +
  geom_errorbar(aes(ymin = lower_095, ymax = upper_095), 
                position = position_dodge(width = 0.6), 
                width = 0.2) +
  geom_hline(yintercept = 1, linetype = "dashed", color = "gray40") +
  coord_flip() +
  scale_y_log10() +
  labs(
    title = "Hazard Ratios for Death and Recurrence",
    x = "",
    y = "Hazard Ratio (log scale)",
    color = "Outcome"
  ) +
  theme_minimal()

forestplot_recurrence_death_Pattern


#With pvalues
combined_table_Pattern$p_label <- ifelse(
  combined_table_Pattern$pval < 0.001, 
  "p < 0.001", 
  paste0("p = ", formatC(combined_table_Pattern$pval, format = "f", digits = 3))
)



forestplot_recurrence_death_Pattern_pvalues <- ggplot(combined_table_Pattern, aes(x = reorder(Variable, -HR), y = HR, color = Outcome)) +
  geom_point(position = position_dodge(width = 0.6), size = 3) +
  geom_errorbar(aes(ymin = lower_095, ymax = upper_095), 
                position = position_dodge(width = 0.6), 
                width = 0.2) +
  geom_text(aes(label = p_label), 
            position = position_dodge(width = 0.6), 
            hjust = -0.2, 
            size = 3) +  # Adjust size and hjust as needed
  geom_hline(yintercept = 1, linetype = "dashed", color = "gray40") +
  coord_flip() +
  scale_y_log10() +
  labs(
    title = "Hazard Ratios for Death and Recurrence",
    x = "",
    y = "Hazard Ratio (log scale)",
    color = "Outcome"
  ) +
  theme_minimal()

forestplot_recurrence_death_Pattern_pvalues

```







#5. Combination HE pattern + HE tumor budding


#Combination HE pattern + HE tumor budding_Kaplan Meyer recurrence individual plot
```{r}
# Create the survival object
surv_obj <- Surv(time = TP53status_clinical$time_recurrence_surv_60m, 
                 event = as.numeric(TP53status_clinical$recurrence_60m == "Yes"))

# Fit the Kaplan-Meier survival model
km_fit <- survfit(surv_obj  ~ Combination2, data = TP53status_clinical)


# Plot the Kaplan-Meier survival curves
p <- ggsurvplot(
  km_fit, 
  data = TP53status_clinical,
  pval = TRUE,                    # Display the p-value for the log-rank test
  title = "Kaplan-Meier Survival Curve by Combination2_Recurrence",
  risk.table = TRUE,              # Include the risk table
  risk.table.height = 0.3,        # Adjust the height of the risk table
  xlab = "Time to Event (Months)",# Label for x-axis
  ylab = "Survival Probability",   # Label for y-axis
  break.time.by = 10,               # Adjust time interval to match plot
  palette = custom_colors, # Use the custom color palette
  legend = c(0.8, 0.2),           # Adjust legend position (optional)
  legend.title = "Clusters"  # Customize legend labels
) 

# Add the title to the plot
p$plot <- p$plot + ggtitle("Kaplan-Meier Survival Curve by Combination2_Recurrence")

# Display the plot
print(p)


```


#Combination HE pattern + HE tumor budding_Kaplan Meyer death individual plot 
```{r}
# Create the survival object
surv_obj <- Surv(time = TP53status_clinical$time_death_surv_60m, 
                 event = as.numeric(TP53status_clinical$death_disease_60m == "Yes"))

# Fit the Kaplan-Meier survival model
km_fit <- survfit(surv_obj  ~ Combination2, data = TP53status_clinical)


# Plot the Kaplan-Meier survival curves
p <- ggsurvplot(
  km_fit, 
  data = TP53status_clinical,
  pval = TRUE,                    # Display the p-value for the log-rank test
  title = "Kaplan-Meier Survival Curve by Combination2_Death",
  risk.table = TRUE,              # Include the risk table
  risk.table.height = 0.3,        # Adjust the height of the risk table
  xlab = "Time to Event (Months)",# Label for x-axis
  ylab = "Survival Probability",   # Label for y-axis
  break.time.by = 10,               # Adjust time interval to match plot
  palette = custom_colors, # Use the custom color palette
  legend = c(0.8, 0.2),           # Adjust legend position (optional)
  legend.title = "Clusters"  # Customize legend labels
) 

# Add the title to the plot
p$plot <- p$plot + ggtitle("Kaplan-Meier Survival Curve by Combination2_Death")

# Display the plot
print(p)
```


#Combination HE pattern + HE tumor budding_Cox univariate recurrence
```{r}
univ_recurrence_Combination2 <- coxphf(
  Surv(time_recurrence_surv_60m, recurrence_60m) ~ Combination2,
  data = TP53status_clinical,
  maxstep = 0.0001, 
  maxit = 100000
)

# Extract coefficients, hazard ratios, and confidence intervals
univ_recurrence_Combination2_table <- data.frame(
  coef = univ_recurrence_Combination2$coefficients,
  HR = exp(univ_recurrence_Combination2$coefficients),
  lower_095 = univ_recurrence_Combination2$ci.lower,
  upper_095 = univ_recurrence_Combination2$ci.upper,
  pval = univ_recurrence_Combination2$prob
)

# Display the table
kable(univ_recurrence_Combination2_table, digits = 3)
```


#Combination HE pattern + HE tumor budding_Cox multivariate recurrence
```{r}
# Ensure TP53 is a factor
TP53status_clinical$tp53_status_combined <- factor(TP53status_clinical$tp53_status_combined)

# Set "Wt" as the reference level
TP53status_clinical$tp53_status_combined <- relevel(TP53status_clinical$tp53_status_combined, ref = "Wt")


# We check the distribution across groups
# Ensure they are all factors
TP53status_clinical$Combination2 <- factor(TP53status_clinical$Combination2)
TP53status_clinical$figo_ini_ava <- factor(TP53status_clinical$figo_ini_ava)
TP53status_clinical$tp53_status_combined <- factor(TP53status_clinical$tp53_status_combined)

# Create the 3-way table
table <- table(
  TP53status_clinical$Combination2,
  TP53status_clinical$figo_ini_ava,
  TP53status_clinical$tp53_status_combined
)

# Print table
table

# We see that there are groups with 1 case... so we will have to force the results

#As there are problems with sparse data due to 3 variable combination and small groups we use this other paramenters to obtain results
multiv_recurrence_Combination2 <- coxphf(
  Surv(time_recurrence_surv_60m, recurrence_60m) ~ Combination2 + figo_ini_ava + tp53_status_combined,
  data = TP53status_clinical,
  maxstep = 1e-6,
  maxit = 1e6
)

# Extract coefficients, hazard ratios, and confidence intervals
multiv_recurrence_Combination2_table <- data.frame(
  coef = multiv_recurrence_Combination2$coefficients,
  HR = exp(multiv_recurrence_Combination2$coefficients),
  lower_095 = multiv_recurrence_Combination2$ci.lower,
  upper_095 = multiv_recurrence_Combination2$ci.upper,
  pval = multiv_recurrence_Combination2$prob
)

# Display the table
kable(multiv_recurrence_Combination2_table, digits = 3)
```




#Combination HE pattern + HE tumor budding_Cox univariate death
```{r, echo=FALSE}
multiv_death_Combination2 <- coxphf(
  Surv(time_death_surv_60m, death_disease_60m) ~ Combination2 ,
  data = TP53status_clinical,
  maxstep = 0.0001, 
  maxit = 100000
)

# Extract coefficients, hazard ratios, and confidence intervals
multiv_death_Combination2_table <- data.frame(
  coef = multiv_death_Combination2$coefficients,
  HR = exp(multiv_death_Combination2$coefficients),
  lower_095 = multiv_death_Combination2$ci.lower,
  upper_095 = multiv_death_Combination2$ci.upper,
  pval = multiv_death_Combination2$prob
)

# Display the table
kable(multiv_death_Combination2_table, digits = 3)

```




#Combination HE pattern + HE tumor budding_Cox multivariate death
```{r}
multiv_death_Combination2 <- coxphf(
  Surv(time_death_surv_60m, death_disease_60m) ~ Combination2 + figo_ini_ava + tp53_status_combined,
  data = TP53status_clinical,
  maxstep = 0.0001, 
  maxit = 100000
)

# Extract coefficients, hazard ratios, and confidence intervals
multiv_death_Combination2_table <- data.frame(
  coef = multiv_death_Combination2$coefficients,
  HR = exp(multiv_death_Combination2$coefficients),
  lower_095 = multiv_death_Combination2$ci.lower,
  upper_095 = multiv_death_Combination2$ci.upper,
  pval = multiv_death_Combination2$prob
)

# Display the table
kable(multiv_death_Combination2_table, digits = 3)
```


#We plot death and recurrence in a forest plot
```{r}
multiv_death_Combination2_table$Variable <- rownames(multiv_death_Combination2_table)
multiv_recurrence_Combination2_table$Variable <- rownames(multiv_recurrence_Combination2_table)
multiv_death_Combination2_table$Outcome <- "Death"
multiv_recurrence_Combination2_table$Outcome <- "Recurrence"
combined_table_Combination2 <- rbind(multiv_death_Combination2_table, multiv_recurrence_Combination2_table)
library(ggplot2)

combined_table_Combination2$Variable <- factor(combined_table_Combination2$Variable, levels = c(
  "figo_ini_ava1 (Advanced)",
  "Combination2Spray_highbudding",
  "Combination2Spray_lowbudding",
  "tp53_status_combinedMut"
))

forestplot_recurrence_death_Combination2<-ggplot(combined_table_Combination2, aes(x = reorder(Variable, -HR), y = HR, color = Outcome)) +
  geom_point(position = position_dodge(width = 0.6), size = 3) +
  geom_errorbar(aes(ymin = lower_095, ymax = upper_095), 
                position = position_dodge(width = 0.6), 
                width = 0.2) +
  geom_hline(yintercept = 1, linetype = "dashed", color = "gray40") +
  coord_flip() +
  scale_y_log10() +
  labs(
    title = "Hazard Ratios for Death and Recurrence",
    x = "",
    y = "Hazard Ratio (log scale)",
    color = "Outcome"
  ) +
  theme_minimal()

forestplot_recurrence_death_Combination2


#With pvalues
combined_table_Combination2$p_label <- ifelse(
  combined_table_Combination2$pval < 0.001, 
  "p < 0.001", 
  paste0("p = ", formatC(combined_table_Combination2$pval, format = "f", digits = 3))
)



forestplot_recurrence_death_Combination2_pvalues <- ggplot(combined_table_Combination2, aes(x = reorder(Variable, -HR), y = HR, color = Outcome)) +
  geom_point(position = position_dodge(width = 0.6), size = 3) +
  geom_errorbar(aes(ymin = lower_095, ymax = upper_095), 
                position = position_dodge(width = 0.6), 
                width = 0.2) +
  geom_text(aes(label = p_label), 
            position = position_dodge(width = 0.6), 
            hjust = -0.2, 
            size = 3) +  # Adjust size and hjust as needed
  geom_hline(yintercept = 1, linetype = "dashed", color = "gray40") +
  coord_flip() +
  scale_y_log10() +
  labs(
    title = "Hazard Ratios for Death and Recurrence",
    x = "",
    y = "Hazard Ratio (log scale)",
    color = "Outcome"
  ) +
  theme_minimal()

forestplot_recurrence_death_Combination2_pvalues

```



